Transient Corotating Clumps Around Adolescent Low-Mass Stars: Four Years of Complex Rotators from TESS

### 2023-08-29 paper/ms.tex
Update ms
Co-authored-by: Joshua Winn <jnwinn@princeton.edu>

The "complex rotator" crowd made the case.  I bought it!

%%%%%%%%
% CQVS %
%%%%%%%%

@@

\title{Transient Corotating Clumps Around Adolescent Low-Mass Stars: Four Years of Complex Quasiperiodic Variables from TESS}

@@

\received{---}
\revised{---}
\accepted{---}
\shorttitle{Four Years of CQVs}

@@

% 250 word max!
\begin{abstract}
  Complex quasiperiodic variables (CQVs) are low-mass
  pre-main-sequence stars with nearly periodic optical modulation.
  The modulation is probably caused by eclipsing clumps of dust or gas
  in orbit at the Keplerian corotation radius.  Here, we
  report new CQVs discovered in TESS short-cadence data collected
  between 2018~July and 2022~September.  Our search of
  \nstarssearched\ K and M dwarfs with $T$$<$16 and $d$$<$150\,pc
  yielded \ngoods\ high-quality CQVs.  Most of these discoveries are
  new, and they include the brightest ($T$$\approx$9.5), closest
  ($d$$\approx$20\,pc), and oldest ($\approx$200\,Myr) examples of
  this class of object currently known.  A few objects are outliers:
  LP 12-502 for instance showed a ``dip complex'' with a period and
  duration that were fixed over 1,500 cycles, but in detail, this
  system exhibited between four and eight local minima per cycle, and
  at times also displayed multiple periods simultaneously.  We
  demonstrate that transient corotating material is the most viable
  explanation for this class of object, and present evidence
  supporting the hypothesis that this material is sculpted by stellar
  magnetic fields dominated by high order multipoles.  We expect that
  our sample will facilitate future efforts aimed at connecting CQVs
  to the broader contexts of star, disk, and perhaps even exoplanet
  evolution. 	
\end{abstract}

@@

Data from K2 \citep{2014PASP..126..398H} and TESS
\citep{2015JATIS...1a4003R} have revealed a new class of variable
star for which the root cause of variability is only beginning to
become clear: complex quasiperiodic variables (CQVs).  These objects
are identified from their optical light curves, which show nearly
periodic troughs that are either sharp or broad; these troughs are
often superposed on quasi-sinusoidal spot-like modulation
\citep{2017AJ....153..152S,2018AJ....155...63S,2019ApJ...876..127Z}.
Some CQVs show up to eight dips per cycle.  Most CQVs are
pre-main-sequence M dwarfs with ages of $\approx$5-150 million years
(Myr), and rotation periods of 0.2--2 days.  They are observed to
comprise $\approx$1-3\% of M dwarfs younger than 100 Myr
\citep{2016AJ....152..114R,2022AJ....163..144G}.  They generally do
not show near-infrared excesses indicative of dusty disks, but the
wavelength-dependent dip amplitudes of some CQVs is consistent with
reddening by dust
\citep{2017PASJ...69L...2O,2020AJ....160...86B,2022AJ....163..144G,2023MNRAS.518.2921K}.
The dip amplitudes and phases usually evolve gradually over tens to
hundreds of cycles, although they have occasionally been observed to
change abruptly within one cycle
\citep[e.g.][]{2017AJ....153..152S,2022ApJ...925...75P,2023ApJ...945..114P}.

The sharp features of CQV light curves can have durations as short as
5\% of the rotation period ($P_{\rm rot}$), which is too short to be
caused by starspots rotating into and out of view.  Starspots produce
flux variations with characteristic timescales of $P_{\rm rot}$ and
$0.5\,P_{\rm rot}$.  With finely-tuned viewing geometries, starspots
can produce dip durations as short as $\approx$$0.2\,P_{\rm rot}$, but
in such cases, limb darkening causes the dip amplitudes to be smaller
than the observed amplitudes of $\sim$1\% (see
\citealt{2017AJ....153..152S}, Figures 37-41).  Thus, a
``starspot-only'' scenario can be ruled out for many CQVs
\citep{2017AJ....153..152S,2019ApJ...876..127Z,2021MNRAS.500.1366K}.
Given that many CQVs cannot be explained by starspots alone, and
working under the assumption that all CQVs share the same basic
physical scenario, we discard the ``starspot-only'' scenario.
Instead, the correct explanation probably involves material above the
stellar surface that is spatially confined with sharp boundaries
\citep[e.g.][]{2017AJ....153..152S,2022AJ....163..144G}.

Figure~\ref{fig:f1} illustrates two proposed models for the extrinsic
material.  The first scenario invokes opaque dust ``clumps'' that
orbit near the Keplerian corotation radius [$R_{\rm cr} =
(GM/\Omega^2)^{1/3}$, where $\Omega = 2\pi/P_{\rm rot}$] and
periodically transit the star
\citep{2017AJ....153..152S,2023MNRAS.518.4734S}.  The second scenario
invokes ``prominences'', long-lived condensations of cool, dense,
marginally-ionized gas that are embedded within the hotter corona and
corotate with the star
\citep{1989MNRAS.238..657C,2019MNRAS.482.2853J,2022MNRAS.514.5465W}.
These hypothetical structures are analogous to quiescent prominences
and filaments seen in the solar corona \citep[see
e.g.][]{2015ASSL..415.....V}, though rather than existing at a
fraction of the stellar radius as in the solar case, they would exist
at distances of a few stellar radii.  A final possibility is that a thin
optically-thick ring obscures a small portion of the stellar
photosphere \citep{2019ApJ...876..127Z}.  Hot spots passing behind such a structure could produce
sudden dips.  We do not favor this scenario due to the issues
discussed in Appendix~\ref{app:ring}.

@@

CQVs remain mysterious because they have been both hard to discover
and hard to characterize.   They are hard to discover because they are
rare: CQVs comprise $\approx$1\% of the youngest $\approx$1\% of M
dwarfs \citep{2018AJ....155..196R}.  Out of the millions of stars
monitored by K2 and TESS, about 50 CQVs have been reported to date
\citep{2016AJ....152..114R,2017AJ....153..152S,2018AJ....155...63S,2019ApJ...876..127Z,2020AJ....160...86B,2022AJ....163..144G,2023ApJ...945..114P}.
They have been hard to characterize because many of the known CQVs are
relatively faint; the initial K2 discoveries
\citep{2016AJ....152..114R,2017AJ....153..152S} were M2-M6 dwarfs at
distances $\gtrsim$100\,pc, yielding optical brightnesses of
$V$$\approx$15.5 to $V$$>$20.  At such magnitudes, high-resolution
time-series spectroscopy is out of reach with current facilities,
despite the potential utility of such observations.

In this work, we aim to find bright and nearby CQVs, since these
objects will be the most amenable to detailed photometric and
spectroscopic analyses.  To do this, we use 120-second cadence data
acquired by TESS between 2018 July and 2022 September (Sectors 1-55;
Cycles 1-4).  We present our search methods in
Section~\ref{sec:methods}, and the resulting CQV catalog in
Section~\ref{sec:results}.  The observed evolution of many CQVs over a
two-year baseline is described in Section~\ref{sec:evoln}, including a
deep-dive into the behavior of an especially interesting object, LP
12-502.  We discuss a few implications in
Section~\ref{sec:discussion}, and conclude in
Section~\ref{sec:conclusion}.

Some comments on nomenclature are needed.  What we are calling CQVs
have also been called ``transient flux dips'', ``persistent flux
dips'', ``scallop shells'', ``batwings'', \citep{2017AJ....153..152S}
``complex rotators'',
\citep{2019ApJ...876..127Z,2022AJ....163..144G,2023ApJ...945..114P},
and ``complex periodic variables'' \citep{2023MNRAS.518.2921K}.  The
CQVs should not be conflated with ``dippers'', which are classical T
Tauri stars with infrared excesses, and which show large-amplitude
variability linked to obscuring inner disk structures and accretion
hot spots \citep{2014AJ....147...82C,2021ApJ...908...16R}.  The
phenomenology and stellar properties of CQVs and dippers are
quite different (though see Sections~\ref{subsec:irexcess}
and~\ref{subsec:discdippers}).  Despite the cost of introducing yet
another naming standard, we prefer to use a single empirically
descriptive name that highlights these objects as variable stars, and
not ``rotators''.  We also prefer a nomenclature that reflects how the
dips and peaks of these stars always appear to be transient 
when observed over timescales exceeding hundreds of cycles.  ``Complex
quasiperiodic variable'' meets these requirements, for the time being.

@@

		\vspace{-0.6cm}
		\subfloat{
			\includegraphics[width=0.66\textwidth]{f1b.pdf}
		}
	\end{center}
		\vspace{-0.3cm}
	\caption{
		{\bf Complex quasiperiodic variables (CQVs)}:
    {\it Top:} Phase-folded TESS light curves of three CQVs.  Each panel shows
    the average of the data accumulated over one month.
    Gray circles are raw 2-minute data; black
    circles are binned to 300 points per cycle.  The periods in hours
    is printed in the
    bottom right corner.  Left-to-right, the objects are LP
    12-502 (TIC 402980664; Sector~19), TIC 94088626 (Sector 10), and
    TIC 425933644 (Sector~28).
    {\it Bottom:} Cartoon explanations for the phenomenon.  The dust
    clump scenario (left) and prominence scenario (right) both invoke
    magnetically-confined corotating material.
	}
	\label{fig:f1}
\end{figure*}

@@

\subsection{CQV discovery}
\label{subsec:discoverymethods}

Previous methods for finding CQVs have included visually examining
stars in young clusters
\citep{2016AJ....152..114R,2017AJ....153..152S} and automatically
flagging rapid rotators with a large number of strong Fourier
harmonics \citep{2019ApJ...876..127Z}.  The latter approach still
requires visual vetting, since ``stars with strong Fourier harmonics''
is a designation that includes objects such as eclipsing binaries or
multiple stars blended into a single photometric aperture.  In this
work, we implemented a new search approach based on counting the
number of sharp local minima in phase-folded light curves, while also
using the Fourier approach.  We applied these two
search techniques independently.

@@

The dip counting technique aims to count sharp local minima in
phase-folded light curves.  CQVs will preferably have at least three
such minima in order to be distinct from false positives such as
synchronized and spotted binaries (``RS CVn'' stars).

@@

To identify local minima near the edges of the phased light curve,
which usually would cover phases $\phi \in [ 0,1 ]$, we performed the
entire procedure over a phase-folded light curve spanning $\phi \in
[-1,2 ]$, by duplicating and concatenating the ordinary phase-folded
light curve.  The free parameters we adopted throughout this analysis
procedure, for instance the maximum number of spline knots per cycle,
and how large and wide of a local minimum to consider a ``true dip'',
were chosen during testing based on their ability to correctly
re-identify a large fraction ($>$90\%) of previously known CQVs, while also being
able to consistently reject common false positives such as rapidly
rotating spot-induced variability and typical eclipsing binaries.

@@

If a peak of such significance was found, we generated a ``summary
sheet'' with information about the star, its full and folded light
curves, Fourier transform, potential contaminating stars, and
information about these contaminating stars.  In the sectors for which
we performed this analysis, we found that ``summary'' sheets were
generated for $\approx$10\% of the 20{,}000 120-second targets in a
given sector.  We then manually reviewed each summary sheet, and
visually labeled each star based on its morphology (including e.g.
eclipsing binary, CQV, RS CVn, or cataclysmic variable).  CQVs that
were observed in multiple sectors had their light curves stitched
together and analyzed for a more accurate period determination.

@@

\begin{figure*}[!tp]
	\begin{center}
		\centering
		\includegraphics[width=0.96\textwidth]{f2.pdf}
		\vspace{-0.3cm}
		\caption{
			{\bf Validation plots used to  label CQVs}.  The complete figure
			set, with one image per sector for each of the \nallcands\
      objects in Table~\ref{tab:thetable}
			is available in the online journal. {\bf For internal collaboration review:
				\url{https://www.dropbox.com/sh/khtwk5a6z0zgrif/AABRG21Ire7VS4BUnaD1hOR8a?dl=0}}.
			Panels are as follows.
			{\it a)}: Phase-folded light curve; gray points are raw 2-minute
			data and black points are binned to 200 points per cycle.
      The adopted period is given in the lower-right corner.
			{\it b)}: Phase-dispersion minimization (PDM) periodogram.
			Dotted lines show up to the 10$^{\rm th}$ harmonic and
			subharmonic.
			{\it c)}: DSS finder chart, with 1- and 2-TESS pixel radius
			circles displayed for scale.
			{\it d)}: Cleaned light curve, binned to 20-minute cadence, in
			Barycentric TESS Julian Date (BTJD).
			{\it e)}: Phase-folded light curve, binned to 100 points per
			cycle.  The gray line denotes the automated spline-fit to the
			wrapped phase-folded light curve, and small gray triangles
			denote automatically identified local minima.
			{\it f)}: Phase-folded light curve at twice the peak period.
			{\it g)}: Phase-folded light curve at half the peak period.
			{\it h)}: Phase-folded time-series within the ``background''
			aperture defined in the SPOC light curves.
			{\it i)}: Phase-folded flux-weighted centroid in the column
			direction.
			{\it j)}: Phase-folded flux-weighted centroid in the row
			direction.
			{\it k)}: Gaia DR2 color--absolute magnitude diagram.     
			The gray background denotes stars within 100\,pc from \citet{2021A&A...649A...6G}.
			{\it l)}: Information from Gaia DR2, TIC8, and the automated
			dip-counting search pipeline.  ``Neighbors'', abbreviated
			``nbhr'', are listed within apparent distances of 2 TESS pixels
			if $\Delta T$$<$2.5.
			{\it m)}: BANYAN-$\Sigma$ v1.2 association probabilities, calculated
			using positions, proper motions, and the parallax.
		}
		\label{fig:vet}
	\end{center}
\end{figure*}

We assessed whether the objects found using the dip-counting
(Section~\ref{subsec:counting}) and Fourier
(Section~\ref{subsec:fourier}) techniques were consistent with
expectations for CQVs by assembling the data shown in
Figure~\ref{fig:vet}.  We labeled a star as a ``good'' CQV if at least
one TESS sector showed what we viewed as the unambiguous signatures of
the class (short period; at least three dips or else otherwise
oddly-shaped dips; relative stability over a timescale of 30\,days).
We independently noted stars that we thought could be CQVs, but that
were more ambiguous.

Broadly speaking, the most common false positive triggers for both the Fourier
and dip-counting techniques were eclipsing binaries, spot-induced
variability from rapid rotators, and variability from neighboring,
off-target stars.  Typical false positive rates from our dip-counting
pipeline were 5:1, with \nuniqdipflagged\ unique stars flagged, and
about 20\% being labeled either ``good'' or ``possible'' CQVs.  The
Fourier analysis was not amenable to calculating a similar
false-positive rate, because it was implemented as a general
variability search; its results are also being used for analyses on
topics other than CQVs.

@@

\paragraph{Ages}
We estimated the stellar ages by making probabilistic spatial and
kinematic associations between the CQVs and known clusters in the
solar neighborhood.  For most stars in our sample, we did this using
BANYAN\,$\Sigma$
\citep{2018ApJ...856...23G}.\footnote{\url{https://github.com/jgagneastro/banyan_sigma},
git commit \texttt{394b486}} This algorithm calculates the probability
that a given star belongs to one of 27 young clusters (or
``associations'') within 150\,pc of the Sun, by modeling the clusters
as multivariate Gaussians in 3-D position and 3-D velocity space.  We
used the Gaia DR2 sky positions, proper motions, and distances to
calculate the membership probabilities.  BANYAN\,$\Sigma$ in turn
analytically marginalizes over the radial velocity dimension.  The
probabilities returned by this procedure are qualitatively useful, but
should be assessed with caution due to the non-Gaussian nature of most
groups within the solar neighborhood \citep[see
e.g.][Figure~10]{2021ApJ...917...23K}.

@@

\begin{figure*}[!tp]
	\begin{center}
		\centering
		\includegraphics[width=0.98\textwidth]{f3.pdf}
    \vspace{-0.3cm}
		\caption{
      {\bf CQVs found in the TESS 2-minute data.}
      Phased TESS light curves over one month are shown for \ngoods\
      CQVs in the high quality sample.  Gray are raw 2-minute data;
      black bins to 300 points per cycle.  Objects are ordered such
      that sources with the most TESS data available are on top (see
      Section~\ref{sec:catalog}).  
      Each panel is labeled by
      the TIC identifier, the TESS sector number, the period in
      hours, and the three-bit binarity flag from Table~\ref{tab:thetable},
      which denotes Gaia DR3 \texttt{radial\_velocity\_error} outliers
      (bit 1), Gaia DR3 \texttt{ruwe} outliers (bit 2), and stars with
      secondary TESS periods (bit 3). 
		}
		\label{fig:cqvs}
	\end{center}
\end{figure*}

@@

\subsection{CQV catalog}
\label{sec:catalog}

Table~\ref{tab:thetable} lists the \nallcands\ objects identified by
our search.  The ``high-quality sample'' includes \ngoods\ objects,
which demonstrated what we viewed as unambiguous characteristics of
the CQV phenomenon in at least one TESS sector.  The classification of
\nmaybes\ CQV candidates was ambiguous, and the \ndebunked\ remaining
objects were false positives that we discuss below.  The
\texttt{quality} column in the table divides the three classes;
additional data from TESS or other instruments could help resolve the
classification of the ambiguous cases.   Of the \ncqvsnodebunked\ CQVs
and candidate CQVs, \nbothdipfourier\ were found using both the
dip-counting and Fourier techniques, \nyesdipnofourier\ were found
using only the dip-counting technique, and \nyesfouriernodip\ were
found using only the Fourier technique.  In the following, we will
restrict our discussion to the high-quality sample, irrespective of
discovery method.

>>>>

%%%%%%%%
% CRS %
%%%%%%%%

@@

\title{Transient Corotating Clumps Around Adolescent Low-Mass Stars: Four Years of Complex Rotators from TESS}

@@

\received{---}
\revised{---}
\accepted{---}
\shorttitle{Four Years of CRs}

@@

% 250 word max!
\begin{abstract}
  %Complex rotators (CRs) are low-mass
  %pre-main-sequence stars with nearly periodic optical modulation.
  %The modulation is probably caused by eclipsing clumps of dust or gas
  %in orbit at the Keplerian corotation radius.  Here, we report new
  %CRs discovered in TESS short-cadence data collected between
  %2018~July and 2022~September.  Our search of \nstarssearched\ K and
  %M dwarfs with $T$$<$16 and $d$$<$150\,pc yielded \ngoods\
  %high-quality CRs.  Most of these discoveries are new, and they
  %include the brightest ($T$$\approx$9.5), closest
  %($d$$\approx$20\,pc), and oldest ($\approx$200\,Myr) examples of
  %this class of object currently known.  A few objects are outliers:
  %LP 12-502 for instance showed a ``dip complex'' with a period and
  %duration that were fixed over 1,500 cycles, but in detail, this
  %system exhibited between four and eight local minima per cycle, and
  %at times also displayed multiple periods simultaneously.  We
  %demonstrate that transient corotating material is the most viable
  %explanation for this class of object, and present evidence
  %supporting the hypothesis that this material is sculpted by stellar
  %magnetic fields dominated by high order multipoles.  We expect that
  %our sample will facilitate future efforts aimed at connecting CRs
  %to the broader contexts of star, disk, and perhaps even exoplanet
  %evolution. 	
  % % JNW alternate:
  Complex rotators (CRs) are low-mass pre-main-sequence stars with
  light curves characterized by periods of 0.2--2 days, amplitudes of
  a few percent, and slowly-evolving sharp features.  The sharp
  features might be caused by eclipsing clumps of dust or gas orbiting
  at the corotation radius.  To improve our understanding of these
  enigmatic stars, we conducted a survey for CRs using TESS
  short-cadence data collected between 2018~July and 2022~September.
  Our search of \nstarssearched\ K and M dwarfs with $T$$<$16 and
  $d$$<$150\,pc yielded \ngoods\ high-quality CRs, most of which are
  new discoveries.  Our sample includes the brightest
  ($T$$\approx$9.5), closest ($d$$\approx$20\,pc), and oldest
  ($\approx$200\,Myr) known CRs.  One of the new discoveries, LP
  12-502, stands out for the complexity of its light curve --- with as
  many as eight flux dips per cycle --- and the detail with which the
  evolution in each dip's amplitude and phase can be tracked over
  1{,}500 cycles.  We argue that corotating gas or dust remains the
  most viable explanation for CRs. The gas or dust is probably
  entrained by the star's magnetic field, and the sharp features may
  result from a multipolar field topology, a hypothesis supported by
  correspondences between the light curves of CRs and of two B stars
  known to have strong multipolar magnetic fields.
\end{abstract}

@@

Data from K2 \citep{2014PASP..126..398H} and TESS
\citep{2015JATIS...1a4003R} have revealed a new class of variable
star for which the root cause of variability is only beginning to
become clear: complex rotators (CRs).  These objects
are identified from their optical light curves, which show nearly
periodic troughs that are either sharp or broad; these troughs are
often superposed on quasi-sinusoidal spot-like modulation
\citep{2017AJ....153..152S,2018AJ....155...63S,2019ApJ...876..127Z}.
Some CRs show up to eight dips per cycle.  Most CRs are
pre-main-sequence M dwarfs with ages of $\approx$5-150 million years
(Myr), and rotation periods of 0.2--2 days.  They are observed to
comprise $\approx$1-3\% of M dwarfs younger than 100 Myr
\citep{2016AJ....152..114R,2022AJ....163..144G}.  They generally do
not show near-infrared excesses indicative of dusty disks, but the
wavelength-dependent dip amplitudes of some CRs is consistent with
reddening by dust
\citep{2017PASJ...69L...2O,2020AJ....160...86B,2022AJ....163..144G,2023MNRAS.518.2921K}.
The dip amplitudes and phases usually evolve gradually over tens to
hundreds of cycles, although they have occasionally been observed to
change abruptly within one cycle
\citep[e.g.][]{2017AJ....153..152S,2022ApJ...925...75P,2023ApJ...945..114P}.

The sharp features of CR light curves can have durations as short as
5\% of the rotation period ($P_{\rm rot}$), which is too short to be
caused by starspots rotating into and out of view.  Starspots produce
flux variations with characteristic timescales of $P_{\rm rot}$ and
$0.5\,P_{\rm rot}$.  With finely-tuned viewing geometries, starspots
can produce dip durations as short as $\approx$$0.2\,P_{\rm rot}$, but
in such cases, limb darkening causes the dip amplitudes to be smaller
than the observed amplitudes of $\sim$1\% (see
\citealt{2017AJ....153..152S}, Figures 37-41).  Thus, a
``starspot-only'' scenario can be ruled out for many CRs
\citep{2017AJ....153..152S,2019ApJ...876..127Z,2021MNRAS.500.1366K}.
Given that many CRs cannot be explained by starspots alone, and
working under the assumption that all CRs share the same basic
physical scenario, we discard the ``starspot-only'' scenario.
Instead, the correct explanation probably involves material above the
stellar surface that is spatially confined with sharp boundaries
\citep[e.g.][]{2017AJ....153..152S,2022AJ....163..144G}.

Figure~\ref{fig:f1} illustrates two proposed models for the extrinsic
material.  The first scenario invokes opaque dust ``clumps'' that
orbit near the Keplerian corotation radius [$R_{\rm cr} =
(GM/\Omega^2)^{1/3}$, where $\Omega = 2\pi/P_{\rm rot}$] and
periodically transit the star
\citep{2017AJ....153..152S,2023MNRAS.518.4734S}.  The second scenario
invokes ``prominences'', long-lived condensations of cool, dense,
marginally-ionized gas that are embedded within the hotter corona and
corotate with the star
\citep{1989MNRAS.238..657C,2019MNRAS.482.2853J,2022MNRAS.514.5465W}.
These hypothetical structures are analogous to quiescent prominences
and filaments seen in the solar corona \citep[see
e.g.][]{2015ASSL..415.....V}, though rather than existing at a
fraction of the stellar radius as in the solar case, they would exist
at distances of a few stellar radii.  A final possibility is that a
thin optically-thick ring obscures a small portion of the stellar
photosphere \citep{2019ApJ...876..127Z}.  Hot spots passing behind
such a ring could produce sudden dips.  We do not favor this scenario
due to two issues discussed in Appendix~\ref{app:ring}.

@@

CRs remain mysterious because they have been both hard to discover and
hard to characterize.   They are hard to discover because they are
rare: CRs comprise $\approx$1\% of the youngest $\approx$1\% of M
dwarfs \citep{2018AJ....155..196R}.  Out of the millions of stars
monitored by K2 and TESS, about 50 CRs have been reported to date
\citep{2016AJ....152..114R,2017AJ....153..152S,2018AJ....155...63S,2019ApJ...876..127Z,2020AJ....160...86B,2022AJ....163..144G,2023ApJ...945..114P}.
They have been hard to characterize because many of the known CRs are
relatively faint; the initial K2 discoveries
\citep{2016AJ....152..114R,2017AJ....153..152S} were M2-M6 dwarfs at
distances $\gtrsim$100\,pc, yielding optical brightnesses of
$V$$\approx$15.5 to $V$$>$20.  At such magnitudes, high-resolution
time-series spectroscopy is out of reach with current facilities,
despite the potential utility of such observations.

In this work, we aim to find bright and nearby CRs, since these
objects will be the most amenable to detailed photometric and
spectroscopic analyses.  To do this, we use 120-second cadence data
acquired by TESS between 2018 July and 2022 September (Sectors 1-55;
Cycles 1-4).  We present our search methods in
Section~\ref{sec:methods}, and the resulting CR catalog in
Section~\ref{sec:results}.  The observed evolution of many CRs over a
two-year baseline is described in Section~\ref{sec:evoln}, including a
deep-dive into the behavior of an especially interesting object, LP
12-502.  We discuss a few implications in
Section~\ref{sec:discussion}, and conclude in
Section~\ref{sec:conclusion}.

Some comments on nomenclature are needed.  What we are calling
``complex rotators''
\citep{2019ApJ...876..127Z,2022AJ....163..144G,2023ApJ...945..114P}
have also been called ``transient flux dips'', ``persistent flux
dips'', and ``scallop shells'' by \citet{2017AJ....153..152S}, or more
generally ``complex periodic variables'' by
\citet{2023MNRAS.518.2921K}.  The CRs should not be conflated with
``dippers'', which are classical T Tauri stars with infrared excesses,
and which show large-amplitude variability linked to obscuring inner
disk structures and accretion hot spots
\citep{2014AJ....147...82C,2021ApJ...908...16R}.  The phenomenology
and stellar properties of CRs and dippers are quite different (though
see Sections~\ref{subsec:irexcess} and~\ref{subsec:discdippers}).
While we considered other possibilities, we ultimately concluded that
``complex rotator'' correctly draws attention to the fact that
rotation will play a fundamental role in explaining these objects, and
that their key defining feature is the complexity of their optical
light curves relative to the quasi-sinusoidal modulations of
starspots.

@@

		\vspace{-0.6cm}
		\subfloat{
			\includegraphics[width=0.66\textwidth]{f1b.pdf}
		}
	\end{center}
		\vspace{-0.3cm}
	\caption{
		{\bf Complex rotators (CRs)}:
    {\it Top:} Phase-folded TESS light curves of three CRs.  Each panel shows
    the average of the data accumulated over one month.
    Gray circles are raw 2-minute data; black
    circles are binned to 300 points per cycle.  The period in hours
    is printed in the
    bottom right corner.  Left-to-right, the objects are LP
    12-502 (TIC 402980664; Sector~19), TIC 94088626 (Sector 10), and
    TIC 425933644 (Sector~28).
    {\it Bottom:} Cartoon explanations for the phenomenon.  The dust
    clump scenario (left) and prominence scenario (right) both invoke
    magnetically-confined corotating material.
	}
	\label{fig:f1}
\end{figure*}

@@

\subsection{CR discovery}
\label{subsec:discoverymethods}

Previous methods for finding CRs have included visually examining
stars in young clusters
\citep{2016AJ....152..114R,2017AJ....153..152S} and automatically
flagging rapid rotators with a large number of strong Fourier
harmonics \citep{2019ApJ...876..127Z}.  The latter approach still
requires visual vetting, since ``stars with strong Fourier harmonics''
is a designation that includes objects such as eclipsing binaries or
multiple stars blended into a single photometric aperture.  In this
work, we implemented a new search approach based on counting the
number of sharp local minima in phase-folded light curves, while also
using the Fourier approach.  We applied these two
search techniques independently.

@@

The dip counting technique aims to count sharp local minima in
phase-folded light curves.  CRs will preferably have at least three
such minima in order to be distinct from false positives such as
synchronized and spotted binaries (``RS CVn'' stars).

@@

To identify local minima near the edges of the phased light curve,
which usually would cover phases $\phi \in [ 0,1 ]$, we performed the
entire procedure over a phase-folded light curve spanning $\phi \in
[-1,2 ]$, by duplicating and concatenating the ordinary phase-folded
light curve.  The free parameters we adopted throughout this analysis
procedure, for instance the maximum number of spline knots per cycle,
and how large and wide of a local minimum to consider a ``true dip'',
were chosen during testing based on their ability to correctly
re-identify a large fraction ($>$90\%) of previously known CRs, while also being
able to consistently reject common false positives such as rapidly
rotating spot-induced variability and typical eclipsing binaries.

@@

If a peak of such significance was found, we generated a ``summary
sheet'' with information about the star, its full and folded light
curves, Fourier transform, potential contaminating stars, and
information about these contaminating stars.  In the sectors for which
we performed this analysis, we found that ``summary'' sheets were
generated for $\approx$10\% of the 20{,}000 120-second targets in a
given sector.  We then manually reviewed each summary sheet, and
visually labeled each star based on its morphology (including e.g.
eclipsing binary, CR, RS CVn, or cataclysmic variable).  CRs that
were observed in multiple sectors had their light curves stitched
together and analyzed for a more accurate period determination.

@@

\begin{figure*}[!tp]
	\begin{center}
		\centering
		\includegraphics[width=0.96\textwidth]{f2.pdf}
		\vspace{-0.3cm}
		\caption{
			{\bf Validation plots used to  label CRs}.  The complete figure
			set, with one image per sector for each of the \nallcands\
      objects in Table~\ref{tab:thetable}
			is available in the online journal. {\bf For internal collaboration review:
				\url{https://www.dropbox.com/sh/khtwk5a6z0zgrif/AABRG21Ire7VS4BUnaD1hOR8a?dl=0}}.
			Panels are as follows.
			{\it a)}: Phase-folded light curve; gray points are raw 2-minute
			data and black points are binned to 200 points per cycle.
      The adopted period is given in the lower-right corner.
			{\it b)}: Phase-dispersion minimization (PDM) periodogram.
			Dotted lines show up to the 10$^{\rm th}$ harmonic and
			subharmonic.
			{\it c)}: DSS finder chart, with 1- and 2-TESS pixel radius
			circles displayed for scale.
			{\it d)}: Cleaned light curve, binned to 20-minute cadence, in
			Barycentric TESS Julian Date (BTJD).
			{\it e)}: Phase-folded light curve, binned to 100 points per
			cycle.  The gray line denotes the automated spline-fit to the
			wrapped phase-folded light curve, and small gray triangles
			denote automatically identified local minima.
			{\it f)}: Phase-folded light curve at twice the peak period.
			{\it g)}: Phase-folded light curve at half the peak period.
			{\it h)}: Phase-folded time-series within the ``background''
			aperture defined in the SPOC light curves.
			{\it i)}: Phase-folded flux-weighted centroid in the column
			direction.
			{\it j)}: Phase-folded flux-weighted centroid in the row
			direction.
			{\it k)}: Gaia DR2 color--absolute magnitude diagram.     
			The gray background denotes stars within 100\,pc from \citet{2021A&A...649A...6G}.
			{\it l)}: Information from Gaia DR2, TIC8, and the automated
			dip-counting search pipeline.  ``Neighbors'', abbreviated
			``nbhr'', are listed within apparent distances of 2 TESS pixels
			if $\Delta T$$<$2.5.
			{\it m)}: BANYAN-$\Sigma$ v1.2 association probabilities, calculated
			using positions, proper motions, and the parallax.
		}
		\label{fig:vet}
	\end{center}
\end{figure*}

We assessed whether the objects found using the dip-counting
(Section~\ref{subsec:counting}) and Fourier
(Section~\ref{subsec:fourier}) techniques were consistent with
expectations for CRs by assembling the data shown in
Figure~\ref{fig:vet}.  We labeled a star as a ``good'' CR if at least
one TESS sector showed what we viewed as the unambiguous signatures of
the class (short period; at least three dips or else otherwise
oddly-shaped dips; relative stability over a timescale of 30\,days).
We independently noted stars that we thought could be CRs, but that
were more ambiguous.

Broadly speaking, the most common false positive triggers for both the Fourier
and dip-counting techniques were eclipsing binaries, spot-induced
variability from rapid rotators, and variability from neighboring,
off-target stars.  Typical false positive rates from our dip-counting
pipeline were 5:1, with \nuniqdipflagged\ unique stars flagged, and
about 20\% being labeled either ``good'' or ``possible'' CRs.  The
Fourier analysis was not amenable to calculating a similar
false-positive rate, because it was implemented as a general
variability search; its results are also being used for analyses on
topics other than CRs.

@@

\paragraph{Ages}
We estimated the stellar ages by making probabilistic spatial and
kinematic associations between the CRs and known clusters in the
solar neighborhood.  For most stars in our sample, we did this using
BANYAN\,$\Sigma$
\citep{2018ApJ...856...23G}.\footnote{\url{https://github.com/jgagneastro/banyan_sigma},
git commit \texttt{394b486}} This algorithm calculates the probability
that a given star belongs to one of 27 young clusters (or
``associations'') within 150\,pc of the Sun, by modeling the clusters
as multivariate Gaussians in 3-D position and 3-D velocity space.  We
used the Gaia DR2 sky positions, proper motions, and distances to
calculate the membership probabilities.  BANYAN\,$\Sigma$ in turn
analytically marginalizes over the radial velocity dimension.  The
probabilities returned by this procedure are qualitatively useful, but
should be assessed with caution due to the non-Gaussian nature of most
groups within the solar neighborhood \citep[see
e.g.][Figure~10]{2021ApJ...917...23K}.

@@

\begin{figure*}[!tp]
	\begin{center}
		\centering
		\includegraphics[width=0.98\textwidth]{f3.pdf}
    \vspace{-0.3cm}
		\caption{
      {\bf CRs found in the TESS 2-minute data.}
      Phased TESS light curves over one month are shown for \ngoods\
      CRs in the high quality sample.  Gray are raw 2-minute data;
      black bins to 300 points per cycle.  Objects are ordered such
      that sources with the most TESS data available are on top (see
      Section~\ref{sec:catalog}).  
      Each panel is labeled by
      the TIC identifier, the TESS sector number, the period in
      hours, and the three-bit binarity flag from Table~\ref{tab:thetable},
      which denotes Gaia DR3 \texttt{radial\_velocity\_error} outliers
      (bit 1), Gaia DR3 \texttt{ruwe} outliers (bit 2), and stars with
      secondary TESS periods (bit 3). 
		}
		\label{fig:cqvs}
	\end{center}
\end{figure*}

@@

\subsection{CR catalog}
\label{sec:catalog}

Table~\ref{tab:thetable} lists the \nallcands\ objects identified by
our search.  The ``high-quality sample'' includes \ngoods\ objects,
which demonstrated what we viewed as unambiguous characteristics of
the CR phenomenon in at least one TESS sector.  The classification of
\nmaybes\ CR candidates was ambiguous, and the \ndebunked\ remaining
objects were false positives that we discuss below.  The
\texttt{quality} column in the table divides the three classes;
additional data from TESS or other instruments could help resolve the
classification of the ambiguous cases.   Of the \ncqvsnodebunked\ CRs
and candidate CRs, \nbothdipfourier\ were found using both the
dip-counting and Fourier techniques, \nyesdipnofourier\ were found
using only the dip-counting technique, and \nyesfouriernodip\ were
found using only the Fourier technique.  In the following, we will
restrict our discussion to the high-quality sample, irrespective of
discovery method.
